---
title: "Social Survey"
author: "Catherine Simpson"
date: "11/5/17"
output:
  html_document: default
---

```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
source("Import_GSS_2014_haven.R")
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->   

####This study deals with questions about possible relationships between gun ownership and race, sex, political views, opinions on capitol punishment and belief in god. The data is from the 2014 general social survey.

The research question is: 
Was there a relationship between gun ownership and race or sex or political views or opinions on capital punishment or belief in God in 2014? 

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

The data for the social survey has 3,842 observations with 12 different variables all from 2014 data. First the data needed to be organized, so the data was trimmed. An already made R code was used from another file. The select function from the dyplr package extracted the variables that will be used for analysis. The mutate function was used to create three variables for education level. The revalue function was used to organize the data in a better fashion in the following ways: 

-The political views responses were extremely liberal, slightly liberal, extremely conservative, slightly conservative, IAP, and dk. IAP and dk are replaced with value NA

-The responses for belief in God were don't believe and no way to find out changed to no and know God exists was changed to yes. The values some higher power, believe sometimes, believe but doubts and dk were changed to maybe. The response IAP was changed to NA 

-The capital punishment responses yes and no stayed the same. The responses IAP and dk were changed to NA. 

Variable analysis: 

-Belief in God is a factor variable with four levels: no, yes, maybe, and NA. 

-Race is a factor variable with three levels: white, black and other. 

-Capital punishment is a factor variable with three levels: na, yes, and no.

-Sex is a factor variable with two levels: male and female. 

-Political views is a factor variable with three levels: liberals, moderates and conservatives.   
 
All the NA responses will be ommitted when looking at the data. The data will be looked at numerically and graphically. The barcharts will be used for graphical and a two-way table for numerical results. The t-test will be used for the inferential results. 

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


Belief in God vs. Gun ownership bar chart: 

```{r}
god_table=xtabs(~god + owngun, data=n2014,exclude = "NA")
barchartGC(god_table,type="percent", main="Belief in God vs. Gun ownership")
```

All groups in the graph have more people that said no to gun ownership. They all relatively have the same amount of people that said yes to gun ownership. 

Race vs. Gun ownership bar chart: 

```{r}
race_table=xtabs(~race + owngun, data=n2014,exclude = "NA")
barchartGC(race_table,type="percent", main="Race vs. Gun ownership")
```

All groups have more people that said no to gun ownership. Other is the group with the highest amount of people that said no to gun ownership. People that identified as white have the highest amount of people that said yes to gun ownership. 

Sex vs. Gun ownership bar chart: 

```{r}
sex_table=xtabs(~sex + owngun, data=n2014,exclude = "NA")
barchartGC(sex_table,type="percent", main="Sex vs. Gun ownership")
```

More females said no to gun ownership than males while more males said yes to gun ownerhsip. For both groups more people said no to gun ownership. 

Political views vs. Gun ownership bar chart: 

```{r}
polviews_table=xtabs(~polviews + owngun, data=n2014,exclude = "NA")
barchartGC(polviews_table,type="percent", main="Political views vs. Gun ownership")
```

For all groups, more people said no to gun ownership. The liberals had the highest amount of people that said no to gun ownership. Conservatives have the highest group of people that said yes to gun ownership. 

Capital punishment vs. Gun ownership bar chart: 

```{r}
cappun_table=xtabs(~cappun + owngun, data=n2014,exclude = "NA")
barchartGC(cappun_table,type="percent", main= "Capital punishment vs. Gun ownership")
```

All groups have more people that do not own guns. The people that oppose capital punishment have a higher amount of people that said no to gun ownership. The people that favor capital punishment have a higher amount of people that own guns. 



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


Belief in God  vs. Gun ownership two-way table: 

```{r}
rowPerc(god_table)
```

For people that don't believe in God, 32.46% own guns while 67.54% don't own guns. 

For the maybe group, 33.33% own guns while 66.67% don't own guns. 

For people that do believe in God, 33.05% own guns while 66.95% don't own guns. 

Race vs. Gun ownership two-way table: 

```{r}
rowPerc(race_table)
```

For people that identify as white, 39.01% own guns while 60.99% don't own guns. 

For people that identify as black, 16.87% own guns while 83.13% don't own guns. 

FOr the other group. 13.93% own guns while 86.07% don't own guns. 

Sex vs. Gun ownership two-way table: 

```{r}
rowPerc(sex_table)
```

For males, 39.76% own guns while 60.24% don't own guns. 

For females, 27.32% own guns while 72.68% don't own guns.

Political views vs. Gun ownership two-way table: 

```{r}
rowPerc(polviews_table)
```

For liberals, 24.52% own guns while 75.48% don't own guns. 

For moderates, 33.20% own guns while 66.80% don't own guns. 

For conservatives, 42.17% own guns while 57.83% don't own guns. 

Capital punishment vs. Gun ownership two-way table: 

```{r}
rowPerc(cappun_table)
```

For people that favor capital punishment, 41.44% own guns while 58.56% don't own guns. 

For people that oppose capital punishment, 20.58% own guns while 79.42% don't own guns. 


# Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

Although there are graphical and numerical results that provide evidence for our research question, it doesn’t provide enough. The chi-square test will provide the last piece of evidence needed to prove if there is in fact a relationship between gun ownership and the variables: race, sex, belief in God, opinion on capital puishment and political views.

Belief in God vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of belief in God.

Ha: Gun ownership and belief in God are not independant (related).

```{r}
god_table=xtabs(~god + owngun, data=n2014,exclude = "NA")
chisqtestGC(god_table)
```

The chi-square statistic is 0.053. The p-value is 0.9739.This means the probablity of getting a sample like this under these conditions is about a 97% chance. Since the p-value is relatively high, we will fail to reject the null hypothesis and conclude that there is no relationship between the belief in God and gun ownership.

Race vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of race. 

Ha: Gun ownership and race are not independant (related).

```{r}
race_table=xtabs(~race + owngun, data=n2014,exclude = "NA")
chisqtestGC(race_table)
```

The chi-square statistic is 6.1542. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between race and gun ownership. For example if you identify as white the greater the chance you will own a gun.

Sex vs. Gun ownership chi-square test:

Hypotheses:
H0: Gun ownership is independant of sex. 

Ha: Sex and Gun ownership are related (not independant).

```{r}
sex_table=xtabs(~sex + owngun, data=n2014,exclude = "NA")
chisqtestGC(sex_table)
```

The chi-square statistic is 34.9146. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between sex and gun ownership. For example if you are male there is a greater chance you will own a gun.

Capital punishment vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of opinion on capital punishment. 

Ha: Opinion on capital punishment and gun ownership are related (not independant).

```{r}
cappun_table=xtabs(~cappun + owngun, data=n2014,exclude = "NA")
chisqtestGC(cappun_table)
```

The chi-square statistic is 85.1602. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between sex and gun ownership. For example if you favor capital punishment there is a greater chance you will own a gun.

Political views vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of political views. 

Ha: Gun ownership and political views are related (not independant).

```{r}
polviews_table=xtabs(~polviews + owngun, data=n2014,exclude = "NA")
chisqtestGC(polviews_table)
```

The chi-square statistic is 42.3848. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between political views and gun ownership. For example if you identify as conservative there is a greater chance you will own a gun.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

After looking at the graphical, numerical and inferential results, we can conclude:
-Belief in God and gun ownership are independant of each other. This means that there is no relationship between the variables.

-Race and gun ownership are related. The white group and other group are more likely to own guns than the black group.

-Sex and gun ownership are related. Males are more likely to own guns than females.

-Opinion on capital punishment and gun ownership are related. People that favor capital punishment are more likley to own guns than people who oppose capital punishment.

-Political views and gun ownership are related. Conservatives are more likely to own guns than the liberals and moderates.

In conclusion, if there was a profile made on people who own guns. They are most likely going to be a conservative, who favors capital punishment, is male and either white or other. If there was a profile made on people who don’t guns, they are most likely to be female, oppose capital punishment, be liberal or moderate and black.

A variable that could affect the results are whether people from this sample grew up around guns or have ever encountered a gun in their life.
