---
title: "Social Survey"
author: "Put your name here!"
date: "put the date here"
output:
  html_document: default
---

```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
source("Import_GSS_2014_haven.R")
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->   

####This study deals with questions about possible relationships between gun ownership and race, sex, political views, opinions on capitol punishment and belief in god. The data is from the 2014 general social survey.



# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

# Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

Although there are graphical and numerical results that provide evidence for our research question, it doesn’t provide enough. The chi-square test will provide the last piece of evidence needed to prove if there is in fact a relationship between gun ownership and the variables: race, sex, belief in God, opinion on capital puishment and political views.

Belief in God vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of belief in God.

Ha: Gun ownership and belief in God are not independant (related).

```{r}
god_table=xtabs(~god + owngun, data=n2014,exclude = "NA")
chisqtestGC(god_table)
```

The chi-square statistic is 0.053. The p-value is 0.9739.This means the probablity of getting a sample like this under these conditions is about a 97% chance. Since the p-value is relatively high, we will fail to reject the null hypothesis and conclude that there is no relationship between the belief in God and gun ownership.

Race vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of race. 

Ha: Gun ownership and race are not independant (related).

```{r}
race_table=xtabs(~race + owngun, data=n2014,exclude = "NA")
chisqtestGC(race_table)
```

The chi-square statistic is 6.1542. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between race and gun ownership. For example if you identify as white the greater the chance you will own a gun.

Sex vs. Gun ownership chi-square test:

Hypotheses:
H0: Gun ownership is independant of sex. 

Ha: Sex and Gun ownership are related (not independant).

```{r}
sex_table=xtabs(~sex + owngun, data=n2014,exclude = "NA")
chisqtestGC(sex_table)
```

The chi-square statistic is 34.9146. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between sex and gun ownership. For example if you are male there is a greater chance you will own a gun.

Capital punishment vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of opinion on capital punishment. 

Ha: Opinion on capital punishment and gun ownership are related (not independant).

```{r}
cappun_table=xtabs(~cappun + owngun, data=n2014,exclude = "NA")
chisqtestGC(cappun_table)
```

The chi-square statistic is 85.1602. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between sex and gun ownership. For example if you favor capital punishment there is a greater chance you will own a gun.

Political views vs.Gun ownership chi-square test:

Hypotheses:

H0: Gun ownership is independant of political views. 

Ha: Gun ownership and political views are related (not independant).

```{r}
polviews_table=xtabs(~polviews + owngun, data=n2014,exclude = "NA")
chisqtestGC(polviews_table)
```

The chi-square statistic is 42.3848. The p-value is 0. This means the probablity of getting a sample like this under these condition is extremely small if the null hypothesis is true. Therfore we reject the null hypothesis which means that there is a relationship between political views and gun ownership. For example if you identify as conservative there is a greater chance you will own a gun.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
